Template for indicator documentation

### 2023-09-27 slitasje.Rmd

@@

@@

The upper reference value is 0% (no *damage*), and the lower reference
value is 100%. Note that 100% _damage_ does not mean that all the area
must be scarred or damaged, but that all hypothetical quadrats around a
point is affected (the variable is frequency-driven, and not
amount-driven). The threshold for good ecosystem condition is set to 10%
damage.

@@

```{r}
naturetypes %>%
  as.tibble() %>%
  group_by(NiN_variable_code)%>%
  summarise(mean = mean(NiN_variable_value, na.rm=T),
            median = median(NiN_variable_value, na.rm=T))
```

The table above shows that the median value for all the variables is zero. The most common % value for each range category should then be the lower limit of that category. For example, when the category is `1/2 - 1` then the % value is most likely 50%, and not 75% which is the center value for that value range. One  issue is when the variable value is 1, because there the lower limit is the same as when the variable value is 0. Here I think I must manually assign it to something a bit higher than 0.

@@

@@

It appears most localities are in good condition. Note also that we can get >100% for the unscaled variable.

@@

This implies that it is impossible to detect or measure a state when it
is in its most degraded form (100% disturbed). This is a problem with
the original data resolution. But I will compensate somewhat for this by
using a non-linear transformation.

>>>>

The indicator is calculated separately for the three ecosystems, but all nature types associated with the same ecosystem are pooled and treated equally. This means we assume the nature types that are mapped and which we have data from are, when pooled, representative for the main ecosystem. This is less true for Naturally open ecosystems than it is for the other two. You can read more about that [here](#naturtype).

@@

Regarding update frequency, the Nature type data set is not replicated, nor will it be in the future. We must therefore trust that the selection criteria for which areas to survey will not change drastically and systematically over time. For example, we are now surveying a lot in pressure areas, meaning lowland areas close to human population centers. If we in the future allocate more time to surveying wilderness areas, then the indicator becomes biased. On the other hand, ANO is replicated every 5 years.

@@

The upper reference value is 0% (no *damage*), and the lower reference
value is 100%. Note that 100% _damage_ does not mean that all the area
must be scarred or damaged, but that all hypothetical quadrats around a
point is affected (the variable is frequency-driven, and not
amount-driven).

The threshold for good ecosystem condition is set to 10%
damage. This is an expert judgement, and a rather weak one at that. 
In the future we should try to empirically validate the dose-response relationship,
for example by comparing different variable scores with erosion rates or the potential for plant growth.
This would need to be done specifically for each nature type.

@@

The table below shows that, if I assume there is some ADSV and therefore exlude case with variable value <1, the median value for all the variables is 1. The most common % value for each range category should then be the lower limit of that category. For example, when the category is `1/2 - 1` then the % value is most likely 50%, and not 75% which is the center value for that value range. One  issue is when the variable value is 1, because there the lower limit is the same as when the variable value is 0. Here I think I must manually assign it to something a bit higher than 0.
```{r}
naturetypes %>%
  as.tibble() %>%
  filter(NiN_variable_value >0) %>%
  group_by(NiN_variable_code)%>%
  summarise(mean = mean(NiN_variable_value, na.rm=T),
            median = median(NiN_variable_value, na.rm=T))
```

@@

This scaling of the variable implies that it very difficult to measure a location when it
is in its most degraded form (100% disturbed), since the highest % value one can use is 75% for PRSL and PRTK and just 50% for the other two. 
This is a problem with the original data resolution. 
But I will compensate somewhat for this by
using a non-linear transformation.

@@

It appears most localities are in good condition. Note also that we theoretically can get >100% for the unscaled variable.

@@
